Safe Transferable Regions

### 2016-10-13 paper/abstract.tex
minor updates

Recent work in high-performance systems written in managed languages
(such as Java or C\#) has shown that garbage-collection can be a
significant performance bottleneck.  A class of these systems, focused
on big-data, create numerous large data structures with well-defined
lifetimes.  In this paper, we present a language and a memory
management scheme based on user-managed memory regions (called
\emph{transferable regions}) that allow programmers to exploit the
knowledge of data structures' lifetimes to achieve significant
performance improvements.

Manual memory management is susceptible to the usual perils of
dangling pointers. A key contribution of this paper is a
refinement-based region type system that ensures the memory safety of
C\# programs in the presence of transferable regions. We complement
our type system with a type inference algorithm that infers principal
region types for first-order programs, and practically useful types
for higher-order programs. This eliminates the need for programmers to
write region annotations on types, while facilitating the reuse of
existing C\# libraries with no modifications. Experiments demonstrate
the practical utility of our approach.

>>>>

% Recent work in high-performance systems written in managed languages
% (such as Java or C\#) has shown that garbage-collection can be a
% significant performance bottleneck.
% A class of these systems, focused on big-data, create numerous large
% data structures with well-defined lifetimes.
%
There is increasing interest in alternative memory management techniques
that strike a balance between the convenience of garbage collection and
the better performance that manual memory management enables, without
compromising safety.
%
In this paper, we present a language and a memory
management scheme based on user-managed memory regions (called
\emph{transferable regions}) that allow programmers to exploit the
knowledge of data structures' lifetimes to improve performance.

Manual memory management is susceptible to the usual perils of
dangling pointers. A key contribution of this paper is a
refinement-based region type system that ensures the memory safety of
C\# programs in the presence of transferable regions.
We complement our type system with a type inference algorithm.
This eliminates the need for programmers to write region annotations on types.
We establish the soundness of the type inference algorithm and
show that the key steps of constraint-generation and constraint-solving
in the type inference algorithm are complete.

% while facilitating the reuse of existing C\# libraries with no modifications.
% Experiments demonstrate the practical utility of our approach.
